### 2016-03-08 paper/paper.tex
added object movement policy section

\documentclass[twoside]{article}
\usepackage{aistats2015}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{float}
\usepackage{subfig}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{amsmath}

@@

\subsubsection{Estimating the Q-Function}
During the simulation we generate sample in the form $(s_i, a_i, r_i, s'_i)$,
where we start in state $s_i$, take action $a_i$ to get to state $s'_i$ and
get reward $r_i$. We also define a feature-based Q-function as $Q(s, a) =
\phi(s,a)^T \theta$ with feature function $\phi(s,a)$ and parameter vector
$\theta$. Details of our state and action representations and rewards and
features we used are discussed in section TODO.

@@

\subsubsection{Actor Critic REPS}
In this step we find a new policy $\pi'$ which optimizes the expected Q-value
as given by our estimated Q-function, but has a limited Kullback-Leibner (KL)
distance to the old policy $\pi$.

@@

\subsubsection{Obtaining a new Policy}
The previous step only gives us desired probabilities
$p(s_i, a_i) = p(s_i) \pi'(a_i|s_i)$ for our samples so in this step we need
to generalize these by fitting a new policy $\bar{\pi}$.

@@

In this paper we use a weighted Gaussian Process (GP) to obtain our new policy
$\bar{\pi}$ by estimating action $a_i$ for state $s_i$ with weight $w_i$. We
also use the sparse version of the GP to reduce computation time.

@@

TODO citation / explain Sparse GP?

@@

Our learning process is implemented in Python and uses a simple kilobot
simulator (see \autoref{sec:simulator}) written by us to generate samples.

\subsection{Maze Solving Policy}
This policy tells us where the object should move given the current object
position and solves the maze (\autoref{fig:simulator}) in this way.

Because this policy is relatively easy to learn and there are specialized maze
solvers which can easily compute an optimal path through any maze we did not
learn this policy and instead focused on the second one, which is much more
interesting. Therefore we manually specified this policy by providing a number
of waypoints the object should pass and giving the next waypoint as the target
position. Once we reach a waypoint we move one to next one.

In our experiments this simple policy is sufficient to move different objects
through a simple maze using our object movement policy. For this reason we will
only discuss the object moving policy in the following sections.

TODO dont learn first part because already done in first paper?
TODO use maze solver or just points list?

@@

Given these parameters the simulator first generates $\nep$ kilobot start
positions $\mathit{start}_i$ equally spaced in a circle of radius $1.5 \cdot r$
around the object which starts in the middle of the screen where $r$ is the
radius or half width of the object depending on the its shape. Then for
$i \in \{1, \dots, \nep\}$ it does the following:

The object's position is reset to the middle of the screen and its angle is set
to 0. After that the light position is set to $\mathit{start}_i$ and the
position of the $\nkb $kilobots is set to $\mathit{start}_i + \mathit{off}_j$,
where $\mathit{off}_j$ is a small individual offset for kilobot number $j$.
Then $\nst$ time steps are simulated where in each one an action $a_k$ is chosen
for the current state $s_k$ according to $\pi$ and applied by adding $a_k$ to
the current light posiiton. Next the kilobots move directly toward the light and
lastly we record the current state as $s_k'$, compute the reward $r_k$ and
save the tuple $(s_k, a_k, r_k, s_k')$ as a sample. We also restrict the
magnitude of the light and kilobot movements to get a more realistic simulation.

@@

As we will discuss in \autoref{sec:objpolicy} our object movement policy learns
how to move the light to push the object to the right given the current light
and kilobot positions relative to the object's position. The maze solving policy
now gives us a target position $t$ from which we compute the target direction as
$d = t - o$ where $o$ is the current object position. We mow compute the angle
$\alpha$ between $d$ and $(1, 0)$ (the vector pointing to the right) and rotate
the state around $\alpha$. Then we use the object movement policy to get an
action for our rotated state which we rotate by $-\alpha$ to get the final
action which we apply just as in the sample generation case.

@@

% sample generation iteration
% states
% features

>>>>

\documentclass[twoside]{article}
\usepackage{aistats2015}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{float}
\usepackage{subfig}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amssymb}

@@

\subsubsection{Estimating the Q-Function}
\label{sec:estimateq}

During the simulation we generate sample in the form $(s_i, a_i, r_i, s'_i)$,
where we start in state $s_i$, take action $a_i$ to get to state $s'_i$ and
get reward $r_i$. We also define a feature-based Q-function as $Q(s, a) =
\phi(s,a)^T \theta$ with feature function $\phi(s,a)$ and parameter vector
$\theta$. Details of our state and action representations and rewards and
features we used are discussed in section TODO.

@@

\subsubsection{Actor Critic REPS}
\label{sec:reps}

In this step we find a new policy $\pi'$ which optimizes the expected Q-value
as given by our estimated Q-function, but has a limited Kullback-Leibner (KL)
distance to the old policy $\pi$.

@@

\subsubsection{Obtaining a new Policy}
\label{sec:gp}

The previous step only gives us desired probabilities
$p(s_i, a_i) = p(s_i) \pi'(a_i|s_i)$ for our samples so in this step we need
to generalize these by fitting a new policy $\bar{\pi}$.

@@

In this paper we use a weighted Gaussian Process (GP) to obtain our new policy
$\bar{\pi}$ by estimating action $a_i$ for state $s_i$ with weight $w_i$. We
also use the sparse version of the GP as introduced in \cite{sparsegp} to reduce
computation time.

@@

@@

Our learning process is implemented in Python and uses a simple kilobot
simulator (see \ref{sec:simulator}) written by us to generate samples.

\subsection{Maze Solving Policy}
This policy tells us where the object should move given the current object
position and solves a maze (\autoref{fig:simulator}) in this way.

Because this policy is relatively easy to learn and there are specialized path
planning algorithms like $A^*$ which can easily compute an optimal path through
any maze we did not learn this policy and instead focused on the second one,
which is much more interesting. Therefore we manually specified this policy by
providing a number of waypoints the object should pass and giving the next
waypoint as the target position. Once we reach a waypoint we move one to next
one.

In our experiments this simple policy is sufficient to move different objects
through a simple maze using our object movement policy. For this reason we will
only discuss the object movement policy in the following sections.

@@

Given these parameters the simulator first generates $\nep$ kilobot start
positions $\mathit{start}_i$ equally spaced in a circle of radius $1.5 \cdot r$
around the object which starts in the middle of the screen where $r$ is the
radius or half width of the object depending on the its shape. Then for
$i \in \{1, \dots, \nep\}$ we do the following:

The object's position is reset to the middle of the screen and its angle is set
to 0. After that the light position is set to $\mathit{start}_i$ and the
position of the $\nkb$ kilobots is set to $\mathit{start}_i + \mathit{off}_j$,
where $\mathit{off}_j$ is a small individual offset for kilobot number $j$.
Then $\nst$ time steps are simulated where in each one an action $a_k$ is chosen
for the current state $s_k$ according to $\pi$ and applied by adding $a_k$ to
the current light positon. Next the kilobots move directly toward the light for
a short amount time (1 second for all of our simulations) and lastly we record
the current state as $s_k'$, compute the reward $r_k$ and save the tuple $(s_k,
a_k, r_k, s_k')$ as a sample. We also restrict the magnitude of the light and
kilobot movements to get a more realistic simulation.

@@

As we will discuss in \ref{sec:objpolicy} our object movement policy learns
how to move the light to push the object to the right given the current light
and kilobot positions relative to the object's position. The maze solving policy
now gives us a target position $t$ from which we compute the target direction as
$d = t - o$ where $o$ is the current object position. We now compute the angle
$\alpha$ between $d$ and $(1, 0)$ (the vector pointing to the right) and rotate
the state around $\alpha$. Then we use the object movement policy to get an
action for our rotated state which we rotate by $-\alpha$ to get the final
action which we apply just as in the sample generation case.

@@

For this policy we decided to only learn how to push the object to the right
instead of an arbitrary direction. This simplifies and speeds up the learning
process because we need to generate fewer samples and can represent our policy
using a smaller state subset for the sparse GP and therefore can compute the
policy update much faster. As discussed in the last paragraph of the previous
section we can still use this policy to push the object in an arbitrary
direction by rotation the state and action appropriately.

\subsubsection{On-Policy Learning}

To learn the object movement policy we use on-policy learning. During this
process we have matrices $S_k, A_k, R_k, S'_k$ with $n$ rows holding the $n$
current samples which are initially empty and a current policy $\pi_k$. Our
first policy $\pi_0$ predicts random actions $(a_x, a_y)$, where both $a_x$ and
$a_y$ are sampled uniformly from $[-0.015, 0.015]$ (1.5 cm).

In each iteration we use the current policy $\pi_k$ to generate $m$ samples
$(s_i, a_i, r_i, s'_i)$ and add them to the current samples giving matrices
$S_{k+1}, A_{k+1}, R_{k+1}, S'_{k+1}$ holding $n + m$ samples. To keep the
computation tractable we then randomly choose $10000$ of the $n + m$ samples
and replace the current sample matrices by matrices filled with the chosen
$10000$ samples (if $(n + m) < 10000$ nothing changes).

Next we estimate the Q-function as described in \ref{sec:estimateq},
compute a sample weighting as in \ref{sec:reps} and lastly fit a new sparse
GP as discussed in \ref{sec:gp}, which then becomes our new policy $\pi_{k+1}$.

We repeat these steps a fixed number of times and manually inspect the learned
policy. If it is not satisfactory (it can not solve the maze) we continue the
learing process.

\subsubsection{States and Actions}

With $n$ kilobots our state tupel is
$$s = (l_x, l_y, k_{1x}, k_{1y}, \dots, k_{nx}, k_{ny}), $$
where $(l_x, l_y)$ is the light position and $(k_{ix}, k_{iy})$ is the position
of the $i-th$ kilobot. All positions are relative to the object's center
position.

Our actions tuple is
$$a = (a_x, a_y),$$
which is the direction the light should move in.

\subsubsection{Rewards}

Our reward is given by
$$ r = (c_x - p_x) - 0.5 \cdot \mathrm{abs} (c_y - p_y), $$
where $(c_x, c_y)$ is the object's current position and $(p_x, p_y)$ is the
object's position from the last time step. Therefore we reward movement to the
right and punish movement to the left and any movement in $y$ direction.

\subsubsection{Features and Kernels}

For our learing process we need a feature function $\phi(s, a)$ over
state-action pairs to represent our Q-function, a feature function $\phi(s)$
over states for REPS and a kernel function $K(s_1, s_2)$ between states for the
GP.

The feature functions are implemented in terms of kernel functions by
choosing a random subset of $N$ samples $(s_i, a_i)$ in each iteration and
computing the feature vectors for a given state $s$ and action $a$ as:
$$
\phi(s, a) = \left(
\begin{array}{c}
    K'((s_1, a_1), (s, a)) \\
    K'((s_2, a_2), (s, a)) \\
    \dots \\
    K'((s_N, a_N), (s, a))
\end{array} \right),
$$
$$
\phi(s) = \left(
\begin{array}{c}
    K(s_1, s) \\
    K(s_2, s) \\
    \dots \\
    K(s_N, s)
\end{array} \right),
$$
where $K'$ is a kernel function similar to $K$ which compares state-action
pairs. For our experiments we used $N = 200$.

Let $s$ be a state with $n$ kilobots with positions $(k_{ix}, k_{iy})$
and $s'$ be a state with $m$ kilobots with positions $(k'_{jx}, k'_{jy})$ and
let $a$ and $a'$ be the corresponding actions.
We define $K(s, s')$ and $K'((s, a), (s', a'))$ as follows:

First we compute the distance between the two kilobot configurations as
\begin{align*}
d_{kb} =
  \frac{1}{n^2} &\sum_{i,j \in \{1,\dots,n\}}
    \bar{K}((k_{ix}, k_{iy}), (k_{jx}, k_{jy})) \\
- \frac{2}{n \cdot m} &\sum_{\substack{i \in \{1,\dots,n\}\\j \in \{1,\dots,m\}}}
    \bar{K}((k_{ix}, k_{iy}), (k'_{jx}, k'_{jy})) \\
+ \frac{1}{m^2} &\sum_{i,j \in \{1,\dots,m\}}
    \bar{K}((k'_{ix}, k'_{iy}), (k'_{jx}, k'_{jy})),
\end{align*}
where $\bar{K}$ is a 2 dimensional squared exponential kernel with
$$
\bar{K}(k_1, k_2) =
    \exp \left(-0.5 (k_1 - k_2)^T
    \mathrm{diag}\left(\frac{1}{\sigma_{kb}^2}\right) (k_1 - k_2)\right)
$$
and $\sigma_{kb} \in \mathbb{R}^2$ is the bandwidth for the kilobot dimensions.
This allows us to compare states with different numbers of kilobots and the
computed distance is independent of the order of the kilobot positions in the
state vectors.

After that we compute the distance for the other dimensions (light positions for
$K$ and light positions and actions for $K'$) as
$$
d_{other} = (x_1 - x_2)^T \mathrm{diag}\left(\frac{1}{\sigma_{x}^2}\right) (x_1 - x_2),
$$
where $x_1$ and $x_2$ are vectors holding the values for all non-kilobot
dimensions and $\sigma_x$ is the non-kilobot bandwith.

We combine these distances to get the kernel value
$$
\exp(-0.5 (w \cdot d_{other} + (1 - w) \cdot d_{kb})),
$$
where $w \in [0,1]$ weighs the importance of the non-kilobot dimensions.

In our experiments we used $w = 0.5$ and recomputed $\sigma_{kb}$ and $\sigma_x$
in each iteration as the square root of the median of the squared distance
between all sample values for each dimension separately. We also scaled the
bandwidths uniformly in each dimension by some manually specified factors.
